Application of integer linear programming to scheduling clinician on-call service in Toronto hospital: case study

### 2021-02-02 paper/paper.practical.tex
Work on practical section

The development of the scheduling solution started from a relaxed version of the problem
(without constraints on consecutive blocks/weekends).
We first investigated a maximum flow formulation for solving the problem.
In this formulation, the problem is represented as a graph $G=(V, E)$ where
the vertices comprise clinicians and blocks/weekends, and the edges comprise
potential assignments of clinician to a block/weekend.
A similar approach using maximum flow has been used in [XX, XX, XX].
The advantages of a maximum flow formulation are in the computational efficiency
of the algorithm and the interpretability of the constraints and solution in terms of the graph [???].
However, a max flow formulation makes it difficult to incorporate new constraints.
For instance, in order to accommodate the NCB and NCW constraints,
additional edges and nodes need to be added to the graph, potentially reducing 
the computational efficiency and interpretability.
As a result, we opted to reformulate the problem as an ILP, in which the additional constraints
can be incorporated more easily.
[...]

>>>>

The development of the scheduling solution started from a relaxed version of the problem
(without constraints on consecutive blocks/weekends).
We first investigated a maximum flow formulation for solving the problem.
In this formulation, the problem is represented as a graph $G=(V, E)$ where
the vertices comprise clinicians and blocks/weekends, and the edges comprise
potential assignments of clinician to a block/weekend.
A similar approach using maximum flow has been used in [XX, XX, XX].
The advantages of a maximum flow formulation are in the computational efficiency
of the algorithm and the interpretability of the constraints and solution in terms of the graph [???].
However, a max flow formulation makes it difficult to incorporate new constraints.
For instance, in order to accommodate the NCB and NCW constraints,
additional edges and nodes need to be added to the graph, potentially reducing 
the computational efficiency and interpretability.
As a result, we opted to reformulate the problem as an ILP, in which the additional constraints
can be incorporated more easily.

The ILP formulation went through several stages of design and tweaking.
For instance, we originally aimed to optimize only the BR and WR constraints.
However, after receiving feedback on some of the test schedules, we learned that 
clinicians would prefer to have some of their weekend assignments closer to the block assignments.
Once we were satisfied with the schedules produced by the algorithm, we targeted the
user interface. Our goal was to develop an easy-to-use interface for the administrative staff to use,
without requiring knowledge of the algorithm. [DL - do we need to elaborate more on this?]
